Geometric variational problems

### 2020-11-16 NSF2020/Description/background.tex
yny Nov 15

\noindent
{\bf Collaborative Research: Mathematical modeling and simulation of
self-assembling amphiphilic particles in solvent} \\
%{\bf {Collaborative Research: Mathematical modeling and coarse-grained simulations of self-assembly of amphiphilic Janus particles in a solvent}}\\
{\em Rolf Ryham (lead PI, Fordham University), Bryan Quaife (PI,
Florida State University), and Young-Nan Young (PI, New Jersey Institute of Technology)}
\section{Background}
\label{sec:background}
The goal of this collaborative proposal is to use mathematical modeling
and numerical simulations to investigate the dynamic self-assembly of
amphiphilic particles interacting via hydrophobic forces
in a solvent. Amphiphilic particles (such as lipid molecules) possess
both hydrophobic and hydrophilic structures. In a viscous solvent they
self-assemble into meso-/macroscopic structures (such as micelles and
bilayers of lipids) to shield their hydrophobic parts from contact with
the solvent (water) molecules.
%
%\subsection{Hydrophobic Forces}
%\label{sec:hydrophobicforce}
%
Such self-assembly of amphiphiles via hydrophobic forces is ubiquitous in biology and biophysics \cite{Israelachvili1954},
%The hydrophobic force is a ubiquitous molecular interaction in biology \cite{Israelachvili1954}, 
and has been a major source of nonspecific interactions between
nanoparticles in soft matter
\cite{Sanchez-IglesiasEtAl2012_ACSNano,AltantzisEtAl2013_PSC,XieYangLuEtAl2020_COCIS}. 
%The proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles so to design smart materials of desirable properties by
%tuning the geometry and properties of the amphiphilic particles.

@@

The HAP formulation is, to our knowledge, the first demonstration of
bilayer self-assembly by a continuum-based interaction
model~\cite{Noguchi2001, Farago2003, Brannigan2006, Brooks2009,
Wang2013}. Our simulations use Janus particles to model lipid
amphiphiles which are popular in material science and physics for
creating functional materials~\cite{Lee2014, Lee2013}. Janus particles
are typically spherical with a biphasic material label on either
hemisphere, endowing the particle with a directional order. We model an
elongated lipid by elliptical particles with the hydrophobic label
defined along the ellipse's axis. Under the hydrophobic force, with
excluded volume, the Janus particles spontaneously merge and realign
to form bilayers. This occurs only as a result of energy minimization
and does not require artificial inputs.

It is worth emphasizing that the HAP model uses only a few parameters:
interfacial tension, decay length, repulsion strength, and particle
shape. For example, an elastic modulus for stretching a vesicle from
micropipette manipulation directly calibrates our interfacial tension
parameter. This is in direct contrast with pair-potential-based
approaches in MD simulations and coarse-grained models where many more
parameters are required~\cite{Varilly2011, Wang2013}.
%MD simulators have also made measurements and lately there is better and better agreement with reality. 
%But even the simplest coarse grained models based on pair potentials for lipids has many more parameters \cite{Varilly2011,Wang2013} .

As a proof of concept, our work has already tested for elastic energies
for bending, stretching, and tilt of the bilayer assembly. The elastic
coefficients derived from the HAP simulations show strikingly positive
agreement with experimentally determined values~\cite{Fu2018_SIAM}.
Encouraged by these results and the hydrodynamic simulations of
bilayer assembly, the PIs propose to extend the HAP model to make direct
comparisons with experimental results \S\ref{subsec:specific_aim_1} and to
establish that the HAP model has the capability to model
flow phenomena across length scales and time scales
\S\ref{subsec:specific_aim_2} using efficient, high-order numerical
methods \S\ref{subsec:specific_aim_3}.

\begin{wrapfigure}[14]{l}{0.35\textwidth}
\centerline{\includegraphics[width=0.35\textwidth]{figures/PW_fig2.pdf}}
  \vspace{-8pt}
  \caption{\label{fig:demixing} \footnotesize An initial assembly of
  small and large particles spontaneously segregates into two smaller
  bodies.}
\end{wrapfigure}
Over the past decades, researchers have used a number of mathematical
tools to simulate vesicles in a shear flow, including lattice
Boltzmann~\cite{KaouiHartingMisbah2011_PRE}, coarse-grained Brownian
dynamics~\cite{NoguchiTakasu2002_BJ}, phase
field~\cite{DuLiuWang2004_JCP,BibenKassnerMisbah2005_PRE}, level
set~\cite{DoyeuxGuyotChabannesEtAl2013_JCAM}, boundary
integral~\cite{Shravan09,Rahimian15}, and immersed boundary
approaches~\cite{KimLai2010_JCP,KimLai2012_PRE,HuLaiSeolEtAl2016_JCP}.
Most of these approaches assume a mathematical surface, whether
implicitly or explicitly, and define an elastic bending energy of the
surface. These vesicle studies built off of numerical methods for
calculating energy minimizing steady equilibrium shapes of lipid bilayer
membranes, vesicles, and red blood cells. These approaches range from
the finite element~\cite{Bartels,Peng13,RyKlYaCo16,Sinha15},
phase field~\cite{Du05,QiangDu08,Lowengrub13}, and immersed
boundary methods~\cite{Hu,Hu13, KimLai2010_JCP}. PI RR and collaborators
led in part the development of phase field functionals of membrane
elastic energy and approaches to coupling membrane elasticity to
fluids~\cite{0951-7715-18-3-016,Du05,DuEuler,QiangDu09}.

>>>>

\noindent
{\bf Collaborative Research: Mathematical modeling and simulation of
self-assembling amphiphilic particles in solvent} \\
%{\bf {Collaborative Research: Mathematical modeling and coarse-grained simulations of self-assembly of amphiphilic Janus particles in a solvent}}\\
{\em Rolf Ryham (lead PI, Fordham University), Bryan Quaife (PI,
Florida State University), and Young-Nan Young (PI, New Jersey Institute of Technology)}
\section{Background}
\label{sec:background}
The goal of this collaborative proposal is to use mathematical modeling
and numerical simulations to investigate the dynamic self-assembly of
amphiphilic particles interacting via hydrophobic forces
in solvent. Amphiphilic particles (such as lipid molecules) possess
both hydrophobic and hydrophilic structures. In a viscous solvent they
self-assemble into meso-/macroscopic structures (such as micelles and
bilayers of lipids) to shield their hydrophobic parts from contact with
the solvent (water) molecules.
%
%\subsection{Hydrophobic Forces}
%\label{sec:hydrophobicforce}
%
Such self-assembly of amphiphiles via hydrophobic forces is ubiquitous in biology and biophysics \cite{Israelachvili1954},
%The hydrophobic force is a ubiquitous molecular interaction in biology \cite{Israelachvili1954}, 
and has been a major source of nonspecific interactions between
nanoparticles in soft matter
\cite{Sanchez-IglesiasEtAl2012_ACSNano,AltantzisEtAl2013_PSC,XieYangLuEtAl2020_COCIS}. 
%The proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles so to design smart materials of desirable properties by
%tuning the geometry and properties of the amphiphilic particles.

@@

The HAP formulation is, to our knowledge, the first demonstration of
bilayer self-assembly by a continuum-based interaction
model~\cite{Noguchi2001, Farago2003, Brannigan2006, Brooks2009,
Wang2013}. Our simulations use Janus particles to model lipid
amphiphiles which are popular in material science and physics for
creating functional materials~\cite{Lee2014, Lee2013}. Janus particles
are typically spherical with a biphasic material label on either
hemisphere, endowing the particle with a directional order. We model an
elongated lipid by elliptical particles with the hydrophobic label
defined along the ellipse's axis. 
\begin{wrapfigure}[14]{l}{0.35\textwidth}
\centerline{\includegraphics[width=0.35\textwidth]{figures/PW_fig2.pdf}}
  \vspace{-8pt}
  \caption{\label{fig:demixing} \footnotesize An initial assembly of
  small and large particles spontaneously segregates into two smaller
  bodies.}
\end{wrapfigure}
Under the hydrophobic force, with
excluded volume, the Janus particles spontaneously merge and realign
to form bilayers. This occurs only as a result of energy minimization
and does not require artificial inputs.
%

%
It is worth emphasizing that the HAP model uses only a few parameters:
interfacial tension, decay length, repulsion strength, and particle
shape. For example, an elastic modulus for stretching a vesicle from
micropipette manipulation directly calibrates our interfacial tension
parameter. This is in direct contrast with pair-potential-based
approaches in MD simulations and coarse-grained models where many more
parameters are required~\cite{Varilly2011, Wang2013}.
%MD simulators have also made measurements and lately there is better and better agreement with reality. 
%But even the simplest coarse grained models based on pair potentials for lipids has many more parameters \cite{Varilly2011,Wang2013} .

%As a proof of concept, our work has already tested for elastic energies
%for bending, stretching, and tilt of the bilayer assembly. The elastic
%coefficients derived from the HAP simulations show strikingly positive
%agreement with experimentally determined values~\cite{Fu2018_SIAM}.
%Encouraged by these results and the hydrodynamic simulations of
%bilayer assembly, the PIs propose to extend the HAP model to make direct
%comparisons with experimental results \S\ref{subsec:specific_aim_1} and to
%establish that the HAP model has the capability to model
%flow phenomena across length scales and time scales
%\S\ref{subsec:specific_aim_2} using efficient, high-order numerical
%methods \S\ref{subsec:specific_aim_3}.

%\begin{wrapfigure}[14]{l}{0.35\textwidth}
%\centerline{\includegraphics[width=0.35\textwidth]{figures/PW_fig2.pdf}}
%  \vspace{-8pt}
%  \caption{\label{fig:demixing} \footnotesize An initial assembly of
%  small and large particles spontaneously segregates into two smaller
%  bodies.}
%\end{wrapfigure}
Over the past decades, researchers have used a number of mathematical
tools to simulate vesicles in a shear flow, including lattice
Boltzmann~\cite{KaouiHartingMisbah2011_PRE}, coarse-grained Brownian
dynamics~\cite{NoguchiTakasu2002_BJ}, phase
field~\cite{DuLiuWang2004_JCP,BibenKassnerMisbah2005_PRE}, level
set~\cite{DoyeuxGuyotChabannesEtAl2013_JCAM}, boundary
integral~\cite{Shravan09,Rahimian15}, and immersed boundary
approaches~\cite{KimLai2010_JCP,KimLai2012_PRE,HuLaiSeolEtAl2016_JCP}.
Most of these approaches assume a mathematical surface, whether
implicitly or explicitly, and define an elastic bending energy of the
surface. These vesicle studies built off of numerical methods for
calculating energy minimizing steady equilibrium shapes of lipid bilayer
membranes, vesicles, and red blood cells. These approaches range from
the finite element~\cite{Bartels,Peng13,RyKlYaCo16,Sinha15},
phase field~\cite{Du05,QiangDu08,Lowengrub13}, and immersed
boundary methods~\cite{Hu,Hu13, KimLai2010_JCP}. PI RR and collaborators
led in part the development of phase field functionals of membrane
elastic energy and approaches to coupling membrane elasticity to
fluids~\cite{0951-7715-18-3-016,Du05,DuEuler,QiangDu09}.
